\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{array}							% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cite}
\usepackage[final]{fixme}
\usepackage{pdfpages}
\usepackage{tabularx}
\usepackage{fancyheadings}
\usepackage{lastpage}
\usepackage{float}
\restylefloat{table}


\parskip 6pt % 1pt = 0.351 mm
\parindent 0pt

%\title{Requirement Engineering Process in AMIDST}
%\author{The handsome AMIDST guys et. al.}
%\date{Latest version, \today}							% Activate to display a given date or no date

\pagestyle{fancy}
\lhead{\tiny FP7-ICT 619209 / AMIDST}
\chead{\tiny Page {\thepage} of \pageref{LastPage} \\}
\rhead{\tiny Public}
\renewcommand{\footrulewidth}{0.4pt}
\cfoot{}



\begin{document}

%\includepdf{AMiDST_Template_Deliverables.pdf}

%\maketitle
%
%\begin{abstract}
%\end{abstract}
%



% Table of contents
\tableofcontents

\newpage

% Document History

\section*{Document history}

\begin{table}[htbp]
  \centering
  \begin{tabularx}{\linewidth}{|p{17mm}|p{17mm}|X|X|}\hline
    {\bf Version} & {\bf Date} & {\bf Author (Unit)} & {\bf Description} \\ \hline \hline
    v0.3 &  &  & First draft finished  \\ 
\hline
  \end{tabularx}
\end{table}

\newpage


% Document History

\section*{Executive summary}
\label{sec:summary}
\addcontentsline{toc}{section}{\nameref{sec:summary}}

The aim of this document is to track the development process of the AMIDST toolbox. 

The document is structured into sections, such that each section includes the required details for a specific use case, i.e., a general description of the purpose of the use case as well as a categorised list of its associated requirements.

We then associate to each use case a list of so-called \textit{\textbf{functionalities}} (included in independent subsections) that are coded in the toolbox to cover the specific use case and their associated requirements. A \textit{functionality}, or also called a \textit{feature}, is a set of java classes that allows to perform a specific task or that defines a coherent concept within the toolbox. For example, the creation and managment of random variables, reading a data set from a file, building and handling dynamic Bayesian networks, the maximum likelihood estimation, etc. The set of \textit{functionalities} identify those key parts of the toolbox that any developer or toolbox user need to understand in order to make a proper use of the AMIDST toolbox.

Another important part of this document is to detail the time line of the toolbox development. Each of the use cases contain an additional subsection named \textit{Time tables} which contain details about the development phase of each \textit{functionality} associated to this use case. Six main phases are identified: design, prototype, code-review, testing, documentation and (first) release. Through this section, we can easily track the evolution of the use case.

In the last section, we detail the class diagram of each package in the AMIDST toolbox. 

%Finally, the document is concluded with an appendix which contains a pdf document generated by the toolbox Doxygen and which contains all the java docs of all the coded java classes. 

We highlight that the contents detailed here can be later modified or re-organized to create the different deliverables and the final user-manual of the AMIDST toolbox. 
 
%\newpage

%\input{allUseCases}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  DATA BASES (DB) %%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{DataBases_DB/UseCase_DB_DataBases}
\include{DataBases_DB/Attributes}
\include{DataBases_DB/DataInstance}
\include{DataBases_DB/DataFileReader}
\include{DataBases_DB/ScalableDataInstanceManagement}
%\include{DataBases_DB/TimeTable-DB}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  BASIC DATA STRUCTURES (BS) %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{BasicStructures_BS/UseCase_BS_BasicStructures}
\input{BasicStructures_BS/StaticVariables}
\input{BasicStructures_BS/DynamicVariables}
\input{BasicStructures_BS/DAG}
\input{BasicStructures_BS/Distributions}
\input{BasicStructures_BS/BayesianNetwork}
\input{BasicStructures_BS/DynamicDAG}
\input{BasicStructures_BS/2TDBN}
%\input{BasicStructures_BS/TimeTable-BS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  HUGIN LINK (HL) %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{HuginLink_HL/UseCase_HL_HuginLink}
\input{HuginLink_HL/ConverterAmidstHugin}
\input{HuginLink_HL/ConverterHuginAmidst}
%\input{HuginLink_HL/TimeTable-HuginLink}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  IMPORTANCE SAMPLING (HL) %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{ImportanceSampling_IS/UseCase_IS_ImportanceSampling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  MAXIMUM LIKELIHOOD (ML) %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{MaximumLikelihood_ML/UseCase_ML_MaximumLikelihood}
\input{MaximumLikelihood_ML/ExponentialFaimilyDistributions}
%\input{MaximumLikelihood_ML/TimeTable-ML}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  VARIATIONAL MESSEAGE PASSING (VMP)%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{VariationalMessagePassing_VMP/UseCase_VMP_VariationalMessagePassing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  EXPECTATION PROPAGATION (EP)%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{ExpectationPropagation_EP/UseCase_EP_ExpectationPropagation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE MAP INFERENCE DETERMINISTIC (DMAP)  %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{MAPInferenceDeterministic_DMAP/UseCase_DMAP_MAPInferenceDeterministic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE VARIATIONAL MAP INFERENCE (VMAP)  %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{VariationalMAPInference_VMAP/UseCase_VMAP_VariationalMAPInference}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  TAN CLASSIFIER (TAN)%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{TANClassifier_TAN/UseCase_TAN_TANClassifier}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  PARALLEL PC (PPC)%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{ParallelPC_PPC/UseCase_PPC_ParallelPC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE  DYNAMIC CLASSIFIERS (DC)%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{DynamicClassifiers_DC/UseCase_DC_DynamicClassifiers}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% USE CASE FEATURE SELECTION (FS)%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{FeatureSelection_FS/UseCase_FS_FeatureSelection}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%% CLASS DIAGRAM %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{ClassDiagrams/classDiagrams}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%--------------------------- DOXYGEN -----------------------------------------------------------------------
% \includepdf{doxygenPDF/refman.pdf}
%------------------------------------------------------------------------------------------------------------------- 

\bibliographystyle{splncs}
%\bibliography{re}

%\appendix
%\section{Code Description (Doxygen Document)}
%\label{sec:form-fram-requ}
%\includepdf[offset=0 1cm,scale=0.85,pages={-},pagecommand={\pagestyle{fancy}}]{doxygenPDF/refman.pdf}


\end{document}  