\newpage
\section{Use Case - Hugin link (HL)}
\label{UseCase:HL}

\begin{description}
\item[Priority:] Must
\item[Deadline:] M15
\item[Responsible:] A. Fern\'andez
\end{description}

%--------------------------------------------------------------------------------------------
\subsubsection*{Description of the use case}
%--------------------------------------------------------------------------------------------

This use case contains all the functionality needed to link the AMIDST toolbox with the HUGIN software. This linkage is addressed by converting Hugin models into AMIDST models, and vice versa. This feature is extremely useful as it allows expanding the testing possibilities of the AMIDST models within a well-stablished platform as Hugin. Also, the link to Hugin can be used for providing some extra functionality to AMIDST that will not be implemented. Finally, the linkage is useful for comparison purposes, i.e., a new inference algorithm implemented in AMIDST could be compared with some state-of-the-art algorithm included in Hugin.

%--------------------------------------------------------------------------------------------
\subsubsection*{Must-Requirements list of the use case}
%--------------------------------------------------------------------------------------------

\begin{enumerate}
\item Bayesian network converter from AMIDST to Hugin format. 
\item Bayesian network converter from Hugin to AMIDST format.
\item Possibility of saving the converted Hugin network in a \texttt{.net} file.
\end{enumerate}

%--------------------------------------------------------------------------------------------
\subsubsection*{Could-Requirements list of the use case}
%--------------------------------------------------------------------------------------------

\begin{enumerate}
\item Converter from AMIDST to HUGIN  and vice versa of some functionality that is not relevant for model representation.
\end{enumerate}

