\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%%% formatting the code
\usepackage{listings}


\input{../settingsTexDoc.tex}


\usepackage{hyperref}

\begin{document}



\section{Dynamic Bayesian Networks: Code Examples}\label{sec:dbns}


\begin{itemize}
	\item \hyperref[sec:dbns:datastreams]{Data Streams}
	\item \hyperref[sec:dbns:variables]{Dynamic Random Variables}
	\item \hyperref[sec:dbns:dbns]{Dynamic Bayesian networks}
	\begin{itemize}
		\item \hyperref[sec:dbns:dbns:creating]{Creating Dynamic Bayesian networks}
		\item \hyperref[sec:dbns:dbns:creatingLatent]{Creating Dynamic Bayesian Networks with Latent Variables}
		\item \hyperref[sec:dbns:dbns:modif]{Modifying Dynamic Bayesian Networks}
	\end{itemize}
	\item \hyperref[sec:dbns:sampling]{Sampling from Dynamic Bayesian Networks}
	\item \hyperref[sec:dbns:inference]{Inference Algorithms for Dynamic Bayesian Networks}
	\begin{itemize}
		\item \hyperref[sec:dbns:inference:dmap]{The Dynamic MAP Inference}
		\item \hyperref[sec:dbns:inference:dvmp]{The Dynamic Variational Message Passing}
		\item \hyperref[sec:dbns:inference:dis]{The Dynamic Importance Sampling}
	\end{itemize}
	\item \hyperref[sec:dbns:learning]{Dynamic Learning Algorithms}
	\begin{itemize}
		\item \hyperref[sec:dbns:learning:ml]{Maximum Likelihood for DBNs}
		\item \hyperref[sec:dbns:learning:sv]{Streaming Variational Bayes for DBNs}
	\end{itemize}
\end{itemize}
%\subsubsection{Creating Dynamic Bayesian networks}\label{sec:dbns:dbns:creating}
%\subsubsection{Creating Dynamic Bayesian Networks with Latent Variables}\label{sec:dbns:dbns:creatingLatent}
%\subsubsection{Modifying Dynamic Bayesian Networks}\label{sec:dbns:dbns:modif}


%\subsection{Sampling from Dynamic Bayesian Networks}\label{sec:dbns:sampling}
%\subsection{Inference Algorithms for Dynamic Bayesian Networks}\label{sec:dbns:inference}
%\subsubsection{The Dynamic MAP Inference}\label{sec:dbns:inference:dmap}
%\subsubsection{The Dynamic Variational Message Passing}\label{sec:dbns:inference:dvmp}
%\subsubsection{The Dynamic Importance Sampling}\label{sec:dbns:inference:dis}
%\subsection{Dynamic Learning Algorithms}\label{sec:dbns:learning}
%\subsubsection{Maximum Likelihood for DBNs}\label{sec:dbns:learning:ml}
%\subsubsection{Streaming Variational Bayes for DBNs}\label{sec:dbns:learning:sv}


%\begin{itemize}
%\item \hyperref[sec:bns:datastreams]{Data Streams}
%\item \hyperref[sec:bns:randomvars]{Data Streams}
%\item \hyperref[sec:bns:models]{Models}
%\begin{itemize}
%\item \hyperref[sec:bns:models:creating]{Creating BNs}
%\item \hyperref[sec:bns:models:creatinglatent]{Creating Bayesian networks with latent variables}
%\item \hyperref[sec:bns:models:modif]{Modifiying Bayesian networks}
%\end{itemize}
%
%	\item \hyperref[sec:bns:io]{Input/Output}
%\begin{itemize}
%	\item \hyperref[sec:bns:io:iods]{I/O of data streams}
%	\item \hyperref[sec:bns:io:iobn]{I/O of BNs}
%\end{itemize}
%
%
%	\item \hyperref[sec:bns:inference]{Inference}
%\begin{itemize}
%
%	\item \hyperref[sec:bns:inference:engine]{The inference engine}
%	\item \hyperref[sec:bns:inference:vmp]{Variational Message Passing}
%	\item \hyperref[sec:bns:inference:sampling]{Importance Sampling}
%\end{itemize}
%
%\item \hyperref[sec:bns:learning]{Learning Algorithms}
%\begin{itemize}
%	\item \hyperref[sec:bns:learning:batchml]{Maximum Likelihood}
%	\item \hyperref[sec:bns:learning:parallelml]{Parallel Maximum Likelihood}
%	\item \hyperref[sec:bns:learning:svb]{Streaming Variational Bayes}
%	\item \hyperref[sec:bns:learning:parallelsvb]{Parallel Streaming Variational Bayes}
%\end{itemize}
%
%
%
%
%\item \hyperref[sec:bns:conceptdrift]{Concept Drift Methods}
%\begin{itemize}
%	\item \hyperref[sec:bns:conceptdrift:nbayes]{Naive Bayes with Virtual Concept Drift Detection}
%\end{itemize}
%
%\item \hyperref[sec:bns:huginlink]{HuginLink}
%\begin{itemize}
%	\item \hyperref[sec:bns:huginlink:conversion]{Models conversion between AMiDST and Hugin}
%	\item \hyperref[sec:bns:huginlink:io]{I/O of Bayesian Networks with Hugin net format}
%	\item \hyperref[sec:bns:huginlink:inference]{Invoking Hugin's inference engine}
%	\item \hyperref[sec:bns:huginlink:huginTAN]{Invoking Hugin's Parallel TAN}
%\end{itemize}
%
%
%
%\item \hyperref[sec:bns:moalink]{MoaLink}
%\begin{itemize}
%	\item \hyperref[sec:bns:moalink:moaclass]{AMIDST Classifiers from MOA}
%	\item \hyperref[sec:bns:moalink:moareg]{AMIDST Classifiers from MOA}
%\end{itemize}
%
%\end{itemize}


\subsection{Data Streams}\label{sec:dbns:datastreams}

In this example we show how to use the main features of a \textit{DataStream} object. More precisely, we show how to load a dynamic data stream and how to iterate over the \textit{DynamicDataInstance} objects.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/datastream/DataStreamsExample.java}

\subsection{Dynamic Random Variables}\label{sec:dbns:variables}

This example show the basic functionalities related to dynamic variables.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/variables/DynamicVariablesExample.java}

\subsection{Dynamic Bayesian networks}\label{sec:dbns:dbns}
\subsubsection{Creating Dynamic Bayesian networks}\label{sec:dbns:dbns:creating}
This example creates a dynamic BN, from a dynamic data stream, with randomly generated probability distributions, then saves it to a file.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/models/CreatingDBNs.java}


\subsubsection{Creating Dynamic Bayesian Networks with Latent Variables}\label{sec:dbns:dbns:creatingLatent}
This example shows how to create a BN model with hidden variables. We simply create a BN for clustering, i.e., a naive Bayes like structure with a single hidden variable acting as parant of all the remaining observable variables.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/models/CreatingDBNsWithLatentVariables.java}



\subsubsection{Modifying Dynamic Bayesian Networks}\label{sec:dbns:dbns:modif}
This example shows how to create a BN model with hidden variables. We This example shows how to access and modify the conditional probabilities of a Dynamic Bayesian network model.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/models/ModifyingDBNs.java}


\subsection{Sampling from Dynamic Bayesian Networks}\label{sec:dbns:sampling}
This example shows how to use the \textit{DynamicBayesianNetworkSampler} class to randomly generate a dynamic data stream from a given Dynamic Bayesian network.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/utils/DynamicBayesianNetworkSamplerExample.java}


\subsection{Inference Algorithms for Dynamic Bayesian Networks}\label{sec:dbns:inference}


\subsubsection{The Dynamic MAP Inference}\label{sec:dbns:inference:dmap}

This example shows how to use the Dynamic MAP Inference algorithm.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/inference/DynamicMAPInference.java}


\subsubsection{The Dynamic Variational Message Passing}\label{sec:dbns:inference:dvmp}

This example shows how to use the Factored Frontier algorithm with Variational Message Passing for running inference on dynamic Bayesian networks.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/inference/DynamicVMP_FactoredFrontier.java}



\subsubsection{The Dynamic Importance Sampling}\label{sec:dbns:inference:dis}

This example shows how to use the Factored Frontier algorithm with Importance Sampling for running inference in dynamic Bayesian networks.

\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/inference/DynamicIS_FactoredFrontier.java}


\subsection{Dynamic Learning Algorithms}\label{sec:dbns:learning}
\subsubsection{Maximum Likelihood for DBNs}\label{sec:dbns:learning:ml}

This example shows how to learn the parameters of a dynamic Bayesian network using maximum likelihood from a randomly sampled data stream.



\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/learning/MLforDBNsampling.java}



\subsubsection{Streaming Variational Bayes for DBNs}\label{sec:dbns:learning:sv}

This example shows how to learn the parameters of a dynamic Bayesian network using streaming variational Bayes from a randomly sampled data stream.



\includejavasource{../../../../examples/src/main/java/eu/amidst/dynamic/examples/learning/SVBforDBN.java}




\end{document}