%--------------------------------------------------------------------------------------------------
\section{Introduction}
%--------------------------------------------------------------------------------------------------

This document presents an overview of the software library implementation of the AMIDST toolbox. It covers the methodological developments related to the above-described models and concepts, the data source management functionalities associated with both learning and inference engines, as well as the interface to existing softwares such as HUGIN AMIDST API ensuring the interaction between AMIDST toolbox and HUGIN software. Such interface is extremely useful as it allows to efficiently exploit the existing systems and previously developed algorithms.

The structure of the document is as follows: Section \ref{sec:Preliminaries} briefly introduces the required background and notation, and Section \ref{sec:Content} presents the content of the AMIDST software. Next, Section \ref{sec:DataStructures} gives a general overview of the core components of the framework, including data structures for variables, graphs, Bayesian networks, dynamic Bayesian networks, key distributions such as multinomial and conditional linear Gaussian distributions represented in both standard form and as exponential families. Section \ref{sec:DataBases} provides a description of the considered database functionalities that will be used by AMIDST learning and inference algorithms. Section \ref{sec:HuginLink} presents the functionalities defined for transforming AMIDST models to and from HUGIN software. Finally, Section \ref{sec:Conclusion} rounds the document off with conclusions.