\section{Introduction}

Even though the number of algorithms designed for learning on streaming data is increasing, there is still not a unified and well accepted way for evaluating them.  This is because testing and evaluating algorithms that are designed to work on streaming data are generally more complicated than those designed to work on non streaming data.  There are both statistical and computational reasons for this.  

Static data is data, where each instance can be assumed to be identically and independently distributed i.i.d. 
If the data instances in a data stream are i.i.d., then the only new challenge compared to static data is that the data streams are open ended.  It can be seen as a static data that continuously grow in size.  One way to deal with this is to take out a data set of fixed size, a holdout data set, and perform tests on this.  Another way is to let the performance measures be "accumulating" as the data set is growing.  

Various performance measures on data streams, that have been proposed in the literature.  Performance measures involving loss functions on stream data have been proposed in the papers of Gama et. al. \cite{Gam09}, \cite{Gam09_2}, \cite{Gam13}.  The loss function on regression problems is a function of both the predicted value and the ground truth, while the loss function on classification problems is a function of predicted and real class labels.  The holdout error is basically the average loss on a holdout dataset of fixed size.  The predictive sequential, or \emph{prequential} error is defined as the average loss function up to time step $i$, where $i$ is the current time step. 

Stationary data is data that follow a stationary process, i.e. that the joint probability distribution of any time window does not change when shifted in time. Consequently, parameters such as the mean and variance, if they are present, also do not change over time and do not follow any trends.  It was suggested to use  \emph{prequential} error on stationary streams in \cite{Gam13}.  Moreover, in the case when data was slowly drifting, it was suggested to use the prequental error measures with forgetting mechanisms. The forgetting mechanisms involves either using a time window or fading factors.  In paper \cite{Gam13}, convergence towards the Bayes error is shown for all these performance measures provided that the learners are consistent and data is i.i.d.  Moreover, it is shown that in the case of concept-drift, i.e. when data is drifting and learners are changing over time, the prequental error measures with forgetting mechanisms are favorable.

In this paper, we focus the exposition on binary classification, although we are aware that much can be generalized to multi class classification or regression.  Binary classification becomes complex when there is class \emph{imbalance}, as known  on non streaming data as well.   Specialized learning algorithms on imbalanced streams are proposed in \cite{Dit13, Hoe12, Lic10}, where these papers points out that this problem is particularly difficult and effective algorithms for evaluating such a classifier is vital.  In the papers \cite{Dit13, Lic10}, the area under the receiver operator characteristics curve (AUC) is calculated on limited holdout sets, while in \cite{Hoe12} AUC is calculated on the entire stream.

AUC is a popular method for evaluating classification problems where class imbalance is vital, because it is invariant to the class distribution.  AUC calculation requires that the classification rule is based on comparing a score or an output function to a threshold.  In the Amidst software, this condition is often fulfilled as the output of the Bayesian network is commonly a predicted probability of the class. 

AUC has the statistical interpretation that it is the probability that a member of the "positive" class is scored higher than a member of the "negative" class.  AUC is therefore a measure of the \emph{ranking ability} of the output function.  This is particularly relevant if one wants to change the classification threshold as a consequence of changing class distributions or  changing misclassification costs \cite{Wu07}.  Moreover, it also pointed out that AUC is more preferable than accuracy for model evaluation in \cite{Gam10}.  

AUC calculation involves sorting all instances and iteration over the sorted list.  In a streaming context, this means that the calculation is $O(n)$ where $n$ is the length of the data stream.  If the hole stream is used, it may have computational problems related to memory and cpu time.  Moreover, it is not taking into account concept-drift as discussed above.
In \cite{Brz14}, the prequential AUC with a forgetting mechanism is proposed.  The advantages of this measure is that it allows for concept-drift and fixed memory and cpu time constraints.  

In the AMIDST project there are use case scenarios where the data is multiple i.i.d. streams.  These use case scenarios are dealt with by dividing these streams into sub streams over a particular time window, where a evaluation time is in between the start time and the end time.  The data from the start time to the evaluation time is basically the explanatory variables and the data at the end time is basically the class label.  This problem reduces to a problem of non streaming nature, and the evaluation is not taking concept-drift into account.  However, it is worth noting that this problem can be expanded in the sense of prequental loss as of \cite{Gam13} or prequental AUC as of \cite{Brz14}.  

Moreover, there are also use case scenarios in the Amidst project that do not directly fit the prequental performance measures as of \cite{Gam13} and \cite{Brz14}.  All of these measures involve calculating a loss or a score at each time step.
In this report, we will describe a use case scenario where it is important to predict that certain event is happening somewhere inside an \emph{interval} prior to the event actually takes place.  

This problem has been solved by chopping the streams into a number of sub streams that are approximately i.i.d and labeled as either a positive or a negative.  It is important that enough space between sub streams are left so that dependencies are withdrawn.  In essence, the i.i.d. assumption of the sub streams allows one to investigate the output function independently from the prior distributions and the choice of loss functions.  

%To make this a bit more clear, a stream in this context is referred to as an ordered list of elements at each time step.  A sub %stream is a stream on a fixed time window.

%We have generally two approaches for generating the sub streams.  When the data is multiple i.i.d. streams, the i.i.d. condition of the sub streams is ensured by sampling at the exact same window.  As an example, the data for a client in a bank can be described by a data stream and all the clients are described by multiple streams.  The sub streams are all the clients observed over a fixed time window.
%
%A second approach for generating the sub streams, is when data is essentially one stream, but we can assume that sub streams that are far apart from each other can be assumed to be i.i.d. 


%
%It is also important to note that the algorithms themselves are often designed to weight measurements that are close to the actual time step higher than measurements that are further back.  On locally dependent streams, we must therefore not only assume that data are generated from underlying distributions that are time dependent, but also that the algorithms themselves are time dependent.  
%
%Computational challenges are related to the fact that the data come from an open-ended data stream, conceptually infinitely long, which imposes practical challenges related to restrictions on cpu-time and memory allocation.  
%
%\todo{Elaborate more on computational challenges.}

In section \ref{sec:methodology}, AMIDST relevant methodologies for evaluation of both batch and streaming algorithms are identified and discussed.  This section forms the foundation of the next three sections, where the exact evaluation routines for each use case provider is given. These sections contains a description of the requirements related to evaluation as described in Delivery 1.2 and a short description of the algorithms and the data.  At the end of these sections, the methods for evaluating predictive and runtime performances are exposed and discussed.  Section \ref{sec:conclusion} concludes the report.

%
%The test and evaluation procedure includes specification of what metrics are relevant to use to quantify the ability of the AMIDST system, such as loss functions, maximum response-times, memory limits and output format.  The paper also includes considerations about quantitative improvements AMIDST should obtain over the state of the art.



%\quote{\emph{Task description: In this task we will establish formal procedures for testing and evaluating the
%    developed models and algorithms. This includes specification of maximum response-times, output format,
%    relevant formalization of loss functions, investigations into what metrics are relevant to use to quantify
%    the ability of the AMIDST system, and considerations about what quantitative improvements AMIDST should
%    obtain over state of the art.}}
%
%
%From Helge's slides at the WP 3 kickoff meeting:
%\begin{itemize}
%\item Massive datasets: find relevant techniques, ensure scalability, etc.
%\item Online evaluation of streams: find relevant techniques, ensure scalability, define behavior in changing environment, etc.
%\item Significance of results, e.g., considering changing environment vs. ``reproducability'', distribution for test-statistic, significance levels/sizes of test-sets, etc.
%\end{itemize}
